Probabilistic Programming

### 2014-04-21 paper/paper.tex
Further discuss vim autocomplete and web interface

We encode the structure of an AST as a probabilistic model. The model is a data
structure that maps node paths to probability distributions of node types or
values. A node path is a list of node types and relationships that encodes the
unique position of a node in a program's AST. The probability distribution for a
node path is the set of possible node types or values that would appear in the
syntax tree following the given path, and the respective probabilities of their
occurrence following the given path.

@@

See figure~\ref{fig:sample-model} in the appendix for a model for the program
``var x = 34;".

%\lstinputlisting[caption=Scheduler, style=json]{../rand.js}

\subsection{Application}

@@

The vim autocomplete feature is limited to generating full programs on a single
line, and requires explicitly triggering the \texttt{completefunc} feature for
code snippets.

\begin{figure}[h]
	\centering
	\begin{tabular}{c c}

		\begin{subfigure}[h]{6.25cm}
			\centering
			\includegraphics[width=1.00\textwidth]{vim1}
			\caption{Inserting a for loop} \label{fig:vim1}
		\end{subfigure}

		\begin{subfigure}[h]{6cm}
			\centering
			\includegraphics[width=1.00\textwidth]{vim2}
			\caption{Inserting an object property} \label{fig:vim2}
		\end{subfigure}
	\end{tabular}

	\caption{Screenshots of vim using autocomplete}
\end{figure}

\subsubsection{Guided IDE}

@@

Design decisions, HCI factors

@@

>>>>

We encode the structure of an AST as a probabilistic model. The model is a data
structure that maps node paths to probability distributions of node types or
values. A node path is a list of node types and relationships that encodes the
unique position of a node in a program's AST. The probability distribution for a
node path is the set of possible node types or values that would appear in the
syntax tree following the given path, and the respective probabilities of their
occurance following the given path.

@@

See figure~\ref{fig:sample-model} in the appendix for a model for the program
``var x = 34;" In this model we see for example that the path
\texttt{[Program, body, VariableDeclaration, declarations]} leads to
\texttt{VariableDeclarator} with probability 1. This indicates that a Program
node whose body is a VariableDeclaration always has a declaration that is of
type VariableDeclarator.

\clearpage
\subsection{Application}

@@

\begin{figure}[h]
	\centering
	\begin{tabular}{c c}
		\begin{subfigure}[h]{6.25cm}
			\centering
			\includegraphics[width=1.00\textwidth]{vim1}
			\caption{Inserting a for loop} \label{fig:vim1}
		\end{subfigure}
		\begin{subfigure}[h]{6cm}
			\centering
			\includegraphics[width=1.00\textwidth]{vim2}
			\caption{Inserting an object property} \label{fig:vim2}
		\end{subfigure}
	\end{tabular}
	\caption{Screenshots of vim using autocomplete of programs}
\end{figure}

The vim autocomplete feature is limited to inserting program snippets on a single
line, and requires the user to explicitly trigger the \texttt{completefunc}
feature to be prompted with code snippets. To further explore using our model
for editing, we created a web-based editor that more deeply integrates Probabilistic
Programming.

\clearpage
\subsubsection{Guided IDE}

The Guided IDE offers users an environment for building a program entirely using
our model, using program snippets generated at the cursor's context as building
blocks. The editor presents a window for the program text, and and list of
suggestion code snippets. The user can move a cursor through the code window at
positions corresponding to boundaries of program nodes. As the user moves the
cursor, the snippet suggestions are regenerated using the cursor's node path, to
continuosly provide suggestions relevant to the current context of the program. When the
user selects a suggestion snippet, its AST is inserted into the program's AST at
the cursor's position in the tree, and the code window is redrawn. The user can
move the cursor within newly inserted blocks, so that the program can grow
deeper. The initial corpus for the IDE's model consists of the JavaScript files
that comprise our project.

@@

%Design decisions, HCI factors

@@

We can use the Guided IDE to generate programs.
